---
title: "615_Project"
author: "Matthew Zebert"
date: "October 4, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("arm","data.table","faraway","foreign","ggplot2","knitr", "dplyr")
```


Step 1: Select data from codebook and create a subset data frame of the desired predictors

```{r, include=FALSE}
#read in csv
full <-  read.csv("recs2015_public_v1.csv")

#rename desired predictors
quant <-  full$NUMFRIG; #Number of refrigerators used
size <- full$SIZRFRI1; #Size of most-used refrigerator
door <- full$TYPERFR1; #Door arrangement of most-used refrigerator
age <- full$AGERFRI1; #Age of most-used refrigerator
ice <- full$ICE; #Through-the-door ice on most-used refrigerator
size2 <- full$SIZRFRI2; #Size of second most-used refrigerator
door2 <- full$TYPERFR2; #Door arrangement of second most-used refrigerator
age2 <- full$AGERFRI2; #Age of second most-used refrigerator
locate <- full$LOCRFRI2; #Location of second most-used refrigerator
recyl <- full$RECYCAPP; #Received free recycling of old appliance or equipment (for example, a refrigerator)"
star <- full$ESFRIG; #Energy Star qualified refrigerator (respondent reported)
area <- full$REGIONC; #Census Region
climate <- full$CLIMATE_REGION_PUB #Building America Climate Zone
own <-  full$KOWNRENT; # Own or rent


#combine into data frame
df <- cbind(quant, size, door, age, ice, size2, door2, age2, locate, recyl, star, area, climate, own)
df.2 <- data.frame(cbind(quant, size, door, age, ice, size2, door2, age2, locate, recyl, star, area, climate, own))
class(df.2)
attach(df.2)

```
Step 2: Begin to explore the data via tables, summaries, and plots to identify any issues or wrangling necessary

```{r, include=FALSE}

#EDA of size
table(size)
head(size)
summary(size)


#assume NA means they don't have a fride 
size[size < 0] <- 0


#EDA of quant and size with plots
boxplot(quant)
table(quant)
hist(quant)
hist(size)
size.jitter <-  jitter(size, factor = 1, amount = NULL)
quant.jitter  <-  jitter(quant, factor = 1, amount = NULL)
plot(quant.jitter, size.jitter)

#EDA of fridge #2
table(size2)
head(size)
summary(size)

#assume that NA means they don't have a second fridge
size2[size2 < 0] <- 0


#Remove NA's from age
age[age < 0] <- NA
age2[age2 < 0] <- NA

age.jitter <-   jitter(age, factor = 1, amount = NULL)
age2.jitter <-   jitter(age2, factor = 1, amount = NULL)
plot(age.jitter, quant.jitter)

#age uses 41 and 42 as responses which throws off the plot
#replace 41, 42, 5 response in age with 4, 5, and 6
age.test <-  age
age.test <- replace(age.test, which(age.test == 41), 4)
age.test <- replace(age.test, which(age.test == 5), 6)
age.test <- replace(age.test, which(age.test == 42), 5)
#df$q2 <- replace(df$q, which(df$q < 0), NA)
table(age.test)
table(age)

age.jitter2 <-   jitter(age.test, factor = 1, amount = NULL)
age2.jitter <-   jitter(age2, factor = 1, amount = NULL)
plot(age.jitter2, quant.jitter)


#how does owning the place affect age and quant
#ggplot(data=wdt, aes(x=dist, y=switch)) 

ggplot(data=df.2, aes(x=quant, y = age.test)) +
  geom_jitter(position = position_jitter(height=.01)) +
  facet_wrap(~ own)

class(df)
```



Step 3: Finalize EDA via tables, plots and summaries to make inferences from. 
Looking at Quant, Size, Age, Age of second fridge initially to understand how the data is distributed 

```{r}

hist(age.test)
# Shows normal distribution of the age of fridges. It might have been assumed that the age was negative exponential, but it is not.

hist(quant)
#The number of the fridges used has an expected distribution

hist(size)
#The size histogram shows that catagory 3 and 4, Medium (17.6 to 22.5 cubic feet) and Large (22.6 to 29.5 cubic feet) respectively

#Looking at how the Age of your first fridge compares to the number of fridges
plot(age.jitter2, quant.jitter)
#The distribution is as expected, showing fewer fridges as the age of the fridges increase


#lastly look at how your housing status of owner, renter, or neither affects the age or quantitiy attributes (1= Owned or being bought by someone in your household, 2= Rented, 3= Occupied without payment of rent)
ggplot(data=df.2, aes(x=quant, y = age.test)) +
  geom_jitter(position = position_jitter(height=.01)) +
  facet_wrap(~ own)

#comparing the 3 plots, it looks that your housing status does not differ in terms of how old or how many refrigerators you have. 

```


factor is for vectors, levels is for Data frames
length, unique, uniqueN, length(unique()) all do similiar counts
table returns frequency table
